---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### Introduction

$H_0$: The vaccine (covid921) IS NOT effective in males 

$H_0$: The vaccine (covid921) IS  effective in males 

##Methods


variable "Treatment" is Categorical

variable "Males" is categorical 

We use helper file cat~cat 

##Methods

##Graphical results 


```{r}
males=filter(FinalData, FinalData$sex=="M")

barchartGC(~treatment + infected,data=males)
barchartGC(~infected +treatment,data=males, type="percent")

```



The graph above shows that more men who took the placebo got covid than those who took the vaccine. 



##Numerical results



```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

The table above shows that 73.8% of males who took the placebo got Covid and that 26.2% 0f men who took the vaccine got Covid. 




## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions