---
title: "Covid91 vaccine study Final2024"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```


A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

## Introduction

$H_0$: The vaccine (covid921) IS NOT effective in males 

$H_A$: The vaccine (covid921) IS  effective in males 

## Methods


variable "Treatment" is Categorical

variable "Males" is categorical 

We use helper file cat~cat 


## Graphical results 


```{r}
males=filter(FinalData, FinalData$sex=="M")

barchartGC(~treatment + infected,data=males)
barchartGC(~infected +treatment,data=males, type="percent")

```



The graph above shows that more men who took the placebo got covid than those who took the vaccine. 



## Numerical results



```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

The table above shows that 73.8% of males who took the placebo got Covid and that 26.2% 0f men who took the vaccine got Covid. 


## Inferential results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The Chi-Squared test above shows that the p-value is .0000000000000002 this is less than 0.05 so we must reject HO. 

The fisher exact test shows that the odds ratio is 2.81|1 this shows that it is effective. 

## Conclusion

In conclusion he covid vaccine IS effective in males this is because 1. the p-value is smaller than 0.05 and 2. the Fisher test showed that the odds ratio is effective (2.81|1) 

# Females

## Introduction

$H_0$: The vaccine (covid921) IS NOT effective in females

$H_A$: The vaccine (covid921) IS  effective in females

## Methods

Variable "Females" is categorical

Variable "Treatment" is categorical 

we use helper file cat~cat

## Descriptive Results 


```{r}
females=filter(FinalData, FinalData$sex=="F")
```

```{r}
barchartGC(~treatment + infected,data=females)
barchartGC(~infected +treatment,data=females, type="percent")

```


The graph shows that more women who took the placebo got covid than women who took the vaccine. 



## Numerical results 


```{r}
table2 <- xtabs(~infected + treatment, data=females)
rowPerc(table2)
colPerc(table2)
```


The table above shows that 64% of women who got the placebo got covid and that 35% of women who got the vaccine got covid. 


## Inferential Results 



```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The chi-squared test shows that the p-value is .00000001 this is less than 0.05 so we must reject HO

the fisher exact test shows that the odds ratio is 1.85|1 this odds ratio is effective.



## Conclusion 

In conclusion the vaccine IS effective in females so  we must REJCT Ho this is because the p-value is less that 0.05 and the odds ratio is shown to be effective ( 1.86|1 )



### LGBTQ

## Introduction 

$H_0$ The Vaccine IS NOT Effective in those of the LGBTQ community 

$H_A$ The Vaccine IS effective in those of the LGBTQ community 


## Methods 

The varriable "LGBTQ" is categorical 

The varriable "treatment" is categorical 

We use helper file cat~cat


## Graphical Results 

```{r}
barchartGC(~treatment + LGBTQ,data=FinalData)
barchartGC(~treatment + LGBTQ,data=FinalData, type="percent")

```

the graph shows that those in the LGBTQ community who took the placebo vs those who took the vaccine was pretty much even 



## Numerical results 

```{r}
table3 <- xtabs(~treatment +LGBTQ, data=FinalData)
rowPerc(table3)
colPerc(table3)
```
The data above shows that the people who are gay and had the placebo (50%) is less than those who had the vaccine (51%)



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The chi squared test shows that the p-value is 0.1099 this is greater than 0.05 therefore we must fail to reject H0. 

The fisher exact test shows that the odds ratio is .94|1 this shows that it is not effective.

## Conclusion 

In conclusion we must fail to reject H0 this means that the covid vaccine is NOT effective in LGBTQ members this is because the p-value was greater than 0.05 (0.1099)  and the odds ratio showed it to be ineffective (.94|1)




### Druggies


## Introduction 


$H_0$ The vaccine is NOT effective in drug users 

$H_A$ The vaccine IS effective in drug users 


## Methods 

variable "Drug User" is categorical

Variable "treatment" is categorical 

We user helper file cat~cat

## Graphical Results 

```{r}
barchartGC(~treatment + DrugUser,data=FinalData)
barchartGC(~treatment + DrugUser,data=FinalData, type="percent")

```


The graph shows that drug users who took the placebo and got covid was pretty much even to those who took the vaccine. 


## Numerical Results 


```{r}
table4 <- xtabs(~treatment + DrugUser, data=FinalData)
rowPerc(table4)
colPerc(table4)
```
The data shows that those who are drug users and took the placebo is 50.16%  and those who are drug users and took the vaccine is 
48.84% so they are about even 
## Inferential results 


```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The chi squared test shows that the p-value is .84 which is greater than 0.05 so we must fail to reject H0

the fisher exact test shows that the odds ratio is 1|1 this shows that the odds are  basically 50/50 in effectiveness this means there is NO relationship between treatment and drug users 



## Conclusion 


In conclusion since the p-value was greater than 0.05 we must fail to reject HO this means that the covid vaccine is NOT effective in drug users. Since the odds ratio was 1|1 this shows that there is basically a 50/50 chance in effectivness. 


# Overall Results and Conclusions